Global code motion of congruent instructions

### 2017-04-18 paper.tex
move punctuation within parentheses

There has been a lot of work both in industry and academia to hoist and sink
code out of branches, and in general global scheduling
\cite{click1995global}. Some relate code-hoisting to code-size optimization
\cite{rosen1988global} and many \cite{barany2013, shobaki2013} use global
scheduling to improve performance. Most of the recent work on global scheduling
are done using integer linear programming (ILP), which results in prohibitively
high compile time. To the best of our knowledge we have not found any reference
which explored \gcm{} of congruent instructions in as much detail us. A part of
our implementation (aggressive code hoisting) is already merged
in \LLVM{} trunk, however, a general implementation of \gcm{} of congruent (non
redundant) expressions is still missing from \GCC{} and \LLVM{} trunk. The main
contributions of this paper are:
\begin{itemize}
\item a new optimistic algorithm to move congruent instructions out of branches;
\item a cost model to reduce interference and hence, reduce spills;
\item a technique to maximize hoisting in an optimistic approach by partitioning
  the list of potential candidates sorted by their DFS visit number;
\item experimental evaluation of our implementation in \LLVM{} which combines
  \SSA{} based liveness analysis, and ranking of expressions to move very busy
  expressions in order to reduce register spills and improve performance in some
  cases.
\end{itemize}

@@

\citet{dhamdhere1988fast} and \citet{steven1997advanced} mention
code hoisting in a data-flow framework. A list of Very Busy Expressions (VBE)
computed which are hoisted in a basic-block where the expression is
anticipable (all the operands are available). This algorithm would hoist as far
as possible without regarding the impact on register pressure and as such a cost
model will be required.  Also the description of VBE is based on the classic
data-flow model and an adaptation to a sparse \SSA{} representation is required.

@@

\subsection{Liveness using DJ-graph}
DJ-graph \cite{Sreedhar1996} is a data structure that augments the dominator
tree with join-edges to keep track of data-flow in a program. We use DJ-graph
and merge-sets to compute liveness of variables as illustrated in
\cite{das2012}.  It is very efficient for computing liveness and does not
require any bitvectors to be maintained for each basic-block. The underlying
simplicity of liveness computation is due to \SSA{} form, where the the values
only flow (from def to use) either through dominator edges or the join edges
(where we insert a PHI). The DJ-graph contains both these edges which allows for
computation of merge-sets for each basic-block, i.e., a set of all basic-blocks
where the values can flow from a particular basic-block. We have implemented the
merge-set computation based on a DJ-graph as illustrated in \cite{das2005}. A
simplified version of the merge-set that we implemented is presented here:

@@

The \GCM{} pass can be broadly divided into the following steps that we will
describe in the rest of this section:
\begin{itemize}
\item find candidates (congruent instructions) suitable for \gcm{} (Section~\ref{subsec:finding-candidates}),
\item compute a point in the program where it is both legal
  (Section~\ref{subsec:legality}) and profitable
  (Section~\ref{subsec:cost-models}) to move the code,
\item move the code to hoist point or sink point
  (Section~\ref{subsec:optimistic}), and
\item update data structures to continue iterative \gcm{} (Section~\ref{subsec:optimistic}).
\end{itemize}

@@

The process of computing \GVN{} can be on-demand (as we come across an
instruction) or, precomputed (computing \GVN{} of all the instructions
beforehand). Which process to choose is determined by the scope of code-hoisting
we want to perform. In a pessimistic approach, we want to hoist a limited set of
instructions from the sibling branches as we iterate the DFS tree bottom-up, it
is sufficient to compute \GVN{} values on-demand. Whereas, in the optimistic
approach, as described in Section~\ref{subsec:optimistic}, we want to move as
many instructions as possible, and it would require \GVN{} values to be
precomputed.

@@

\subsubsection{Legality of hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available (or can be
made available by rematerialization), the scalar computations can be hoisted.

@@

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill), but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after \gcm{} is less than before it will
be moved. Essentially, as long as there is one killed operand, code hoisting
will either decrease or preserve the register pressure.  Similarly, code-sinking
will either decrease or preserve the register pressure as long as there is one
operand killed at most.  The following example explains how \gcm{} of
identical computations can reduce the register pressure.

@@

In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximize the total number of hoist instructions.  By sorting
the list of all the candidates in the increasing order of their depth first
search discovery time stamp \cite{clrs} (DFSIn numbers), we make sure that
candidates closer in the list have their common dominator nearby. In
Figure-\ref{fig:dfsin}, if B3, B4, B5, and B6 have identical computations and if
for some reason, they cannot be hoisted at B0, then we partition the set of
hoistable candidates in their DFSIn order. In this case the DFSIn ordering would
be B3, B4, B5, B6 which will allow the instructions in (B3, B4) to be hoisted at
B1 and those in (B5, B6) to be hoisted at B2. Even if instructions in (B3, B4)
could not be hoisted, for some reason, to B1, the instructions in (B5, B6) can
still be hoisted -- if legal and profitable -- to B2.

@@

% TODO: Rewrite the following.
In our current implementation we keep as many candidates in one set as possible
(greedy approach). To start with, first two candidates are tried for
hoistability (see Section~\ref{subsec:optimistic}), if that passes then the next
instruction is tried (if available) and so on until a non-hoistable instruction
is found. The partition starts from that point.  We split the list at a point
where the legality checks fail to hoist subset of candidates which are legal to
hoist and then start finding new hoisting point for the remaining ones.

@@

The amount of hoisting depends on whether we collect \GVN{} of instructions
before finding candidates (optimistic) or, on-demand (pessimistic). It also
depends on the generality of the \GVN{} algorithm as mentioned earlier in
Section~\ref{subsec:finding-candidates}. We have implemented an optimistic
\gcm{} of congruent instructions which uses the liveness analysis as
illustrated by \citet{das2012} and ranking expressions explained by
\citet{rosen1988global}.

\GCM{} basically consists of two parts, i.e., hoisting and sinking. This
implementation only moves congruent instructions. An immediate guarantee of this
approach is gain in code size (the final executable may be of larger size
because of more inlining). The algorithm prefers hoisting to sinking. If the
dependency of a hoistable candidate is in the same basic-block as the candidate,
then the dependency must also be hoistable, otherwise hoisting will be illegal or
would require a complicated code generation to make it legal. The current
algorithm discards cases if the dependency is neither hoistable nor
rematerializable.

@@

Code hoisting opens new opportunities for other hoistable candidates which were
of higher rank (depended on candidates which got hoisted). Ideally we could
iterate on lower ranking expressions first and then proceed to higher ranking
expressions in the same iteration but \LLVM{'}s \GVN{} infrastructure does not
compute equivalence classes in an effective way. We found it simpler to just
recompute the value numbers and start finding hoistable candidates again.

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that are
candidates for \gcm{}, matching the complexity of \PRE{} on \SSA{} form.
The analyses computed for this pass are Global Value Numbering, Computation of
DJ-graph and merge-sets, Marking Barriers, all are linear in number of
instructions in a function. Liveness analysis is expensive and only performed
on-demand for hoistable candidates so it does not affect the overall compilation
time by much. Other analyses like Alias Analysis, \MemorySSA{} and Dominator Tree
are already available in the \LLVM{} pass pipeline. Although we recompute \GVN{}, and
Barriers for each iteration of the \gcm{}, we have not seen significant
increase in the compilation times (see Section~\ref{tab:compile-time}). We have also
provided appropriate compiler
flags to expedite \gcm{} by bailing out with fewer iterations, or skip the
liveness based profitability analysis to aggressively move the code as long as
they are legal.

@@

Overall there was only a minor change in the performance numbers with a few improvements 
(429.mcf, 458.sjeng, etc.) and a few regressions (447.dealII, 453.povray,
etc.). The \gcm{} pass was run twice in the pass pipeline (at -Ofast),
first after EarlyCSE pass, and then after the inliner. Because EarlyCSE removes
local redundancies, \GCM{} would have to analyze less redundant instructions, also
as inliner creates more redundancies, \GCM{} would have more opportunities to move
congruent instructions out of branches.

We see some nice gains in code size at -Os with \gcm{} (403.gcc, 447.dealII, and
483.xalancbmk) as shown in Table~\ref{tab:code-size}.  The code size listed here
is the text size delta (from linux size command) of the final executable for
each benchmark.  On the other hand, some benchmarks like 473.astar increased in
code size by almost 2\%. This is because once the code-size of a function
decreases (due to \GCM{}), it becomes cheaper to inline. As we can see in
Table~\ref{tab:stats}, four more functions got inlined in 473.astar. Except for
447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got
more functions inlined. Also, since the pass runs early, it affects many
optimizations which rely on the number of instructions, length of the use-def
chain, and other metrics.

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|}
      \hline
 \GCM{}  &\multicolumn{2}{|c|}{Total Instructions} &\multicolumn{2}{|c|}{Scalars}  & \multicolumn{2}{|c|}{Loads} & \multicolumn{1}{|c|}{Stores} &Calls   &Instructions    \\\hline
 metric           &  H          &R        &H      &R       &H           &R              &H=R    &H & sunk \\\hline
 400.perlbench	  &  2387	& 2456	  &  855  & 857	   & 1493	& 1560		& 38	& 1	& 4	        \\\hline
 401.bzip2	  &  176	& 300	  &  27	  & 27	   & 149	& 273		& 0	& 0 	& 0		\\\hline
 403.gcc	  &  5670	& 5761	  &  2558 & 2567   & 3109	& 3191		& 3	& 0	& 1	        \\\hline
 429.mcf	  &  42	        & 43	  &  5	  & 5	   & 37 	& 38		& 0	& 0 	& 0		\\\hline
 433.milc	  &  128	& 130	  &  10	  & 10	   & 117	& 119		& 0	& 1	& 0		\\\hline
 444.namd	  &  262	& 264	  &  113  & 113	   & 148	& 150		& 0	& 1	& 0		\\\hline
 445.gobmk	  &  1021	& 1052	  &  312  & 314	   & 705	& 734		& 3	& 1	& 1	        \\\hline
 447.dealII	  &  10455	& 12325	  &  4009 & 4063   & 6394	& 8210		& 20	& 32	& 0	        \\\hline
 450.soplex	  &  2389	& 2731	  &  927  & 1004   & 1437	& 1702		& 14	& 11	& 0	        \\\hline
 453.povray	  &  1967	& 2269	  &  560  & 590	   & 1389	& 1661		& 4	& 14	& 0	        \\\hline
 456.hmmer	  &  408	& 427	  &  82	  & 82	   & 319	& 338		& 1	& 6	& 0	        \\\hline
 458.sjeng	  &  372	& 399	  &  128  & 138	   & 234	& 251		& 1	& 9	& 1	        \\\hline
 462.libquantum	  &  49	        & 49	  &  17	  & 17	   & 32	      	& 32		& 0	& 0	& 0		\\\hline
 464.h264ref	  &  2953	& 3438	  &  712  & 779	   & 2238	& 2656		& 2	& 1	& 1	        \\\hline
 470.lbm	  &  23	        & 23	  &  4	  & 4	   & 19		& 19		& 0	& 0	& 0		\\\hline
 471.omnetpp	  &  391	& 408	  &  92	  & 92	   & 297	& 314		& 0	& 2	& 0		\\\hline
 473.astar	  &  202	& 222	  &  62	  & 64	   & 140	& 158		& 0	& 0	& 0		\\\hline
 482.sphinx3	  &  168	& 172	  &  32	  & 32	   & 136	& 140		& 0	& 0	& 0		\\\hline
 483.xalancbmk	  &  5894	& 6162	  &  2586 & 2660   & 3301	& 3495		& 4	& 3	& 2	        \\\hline
 Grand total      &34957	&38631	  &13091  &13418   &21694	&25041	        &90	&82     &10             \\\hline
    \end{tabular}
  \end{center}
  \caption{\GCM{} stats on SPEC2006 at -Ofast, H(Hoisted), R(Removed); Loads are
    hoisted the most followed by scalars, stores and calls.}
  \label{tab:code-motion-metric}
\end{table}

>>>>

There has been a lot of work both in industry and academia to hoist and sink
code out of branches, and in general global scheduling
\cite{click1995global}. Some relate code-hoisting to code-size optimization
\cite{rosen1988global} and many \cite{barany2013, shobaki2013} use global
scheduling to improve performance. Most of the recent work on global scheduling
are done using integer linear programming (ILP,) which results in prohibitively
high compile time. To the best of our knowledge we have not found any reference
which explored \gcm{} of congruent instructions in as much detail us. A part of
our implementation (aggressive code hoisting) is already merged
in \LLVM{} trunk, however, a general implementation of \gcm{} of congruent (non
redundant) expressions is still missing from \GCC{} and \LLVM{} trunk. The main
contributions of this paper are:
\begin{itemize}
\item a new optimistic algorithm to move congruent instructions out of branches;
\item a cost model to reduce interference and hence, reduce spills;
\item a technique to maximize hoisting in an optimistic approach by partitioning
  the list of potential candidates sorted by their DFS visit number;
\item experimental evaluation of our implementation in \LLVM{} which combines
  \SSA{} based liveness analysis, and ranking of expressions to move very busy
  expressions in order to reduce register spills and improve performance in some
  cases.
\end{itemize}

@@

\citet{dhamdhere1988fast} and \citet{steven1997advanced} mention
code hoisting in a data-flow framework. A list of Very Busy Expressions (VBE)
computed which are hoisted in a basic-block where the expression is
anticipable (all the operands are available.) This algorithm would hoist as far
as possible without regarding the impact on register pressure and as such a cost
model will be required.  Also the description of VBE is based on the classic
data-flow model and an adaptation to a sparse \SSA{} representation is required.

@@

\subsection{Liveness using DJ-graph}
DJ-graph \cite{Sreedhar1996} is a data structure that augments the dominator
tree with join-edges to keep track of data-flow in a program. We use DJ-graph
and merge-sets to compute liveness of variables as illustrated in
\cite{das2012}.  It is very efficient for computing liveness and does not
require any bitvectors to be maintained for each basic-block. The underlying
simplicity of liveness computation is due to \SSA{} form, where the the values
only flow (from def to use) either through dominator edges or the join edges
(where we insert a PHI.) The DJ-graph contains both these edges which allows for
computation of merge-sets for each basic-block, i.e., a set of all basic-blocks
where the values can flow from a particular basic-block. We have implemented the
merge-set computation based on a DJ-graph as illustrated in \cite{das2005}. A
simplified version of the merge-set that we implemented is presented here:

@@

The \GCM{} pass can be broadly divided into the following steps that we will
describe in the rest of this section:
\begin{itemize}
\item find candidates (congruent instructions) suitable for \gcm{}
  (Section~\ref{subsec:finding-candidates},)
\item compute a point in the program where it is both legal
  (Section~\ref{subsec:legality}) and profitable
  (Section~\ref{subsec:cost-models}) to move the code,
\item move the code to hoist point or sink point
  (Section~\ref{subsec:optimistic},) and
\item update data structures to continue iterative \gcm{} (Section~\ref{subsec:optimistic}.)
\end{itemize}

@@

The process of computing \GVN{} can be on-demand (as we come across an
instruction) or, precomputed (computing \GVN{} of all the instructions
beforehand.) Which process to choose is determined by the scope of code-hoisting
we want to perform. In a pessimistic approach, we want to hoist a limited set of
instructions from the sibling branches as we iterate the DFS tree bottom-up, it
is sufficient to compute \GVN{} values on-demand. Whereas, in the optimistic
approach, as described in Section~\ref{subsec:optimistic}, we want to move as
many instructions as possible, and it would require \GVN{} values to be
precomputed.

@@

\subsubsection{Legality of hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available (or can be
made available by rematerialization,) the scalar computations can be hoisted.

@@

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill,) but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after \gcm{} is less than before it will
be moved. Essentially, as long as there is one killed operand, code hoisting
will either decrease or preserve the register pressure.  Similarly, code-sinking
will either decrease or preserve the register pressure as long as there is one
operand killed at most.  The following example explains how \gcm{} of
identical computations can reduce the register pressure.

@@

In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximize the total number of hoist instructions.  By sorting
the list of all the candidates in the increasing order of their depth first
search discovery time stamp \cite{clrs} (DFSIn numbers,) we make sure that
candidates closer in the list have their common dominator nearby. In
Figure-\ref{fig:dfsin}, if B3, B4, B5, and B6 have identical computations and if
for some reason, they cannot be hoisted at B0, then we partition the set of
hoistable candidates in their DFSIn order. In this case the DFSIn ordering would
be B3, B4, B5, B6 which will allow the instructions in (B3, B4) to be hoisted at
B1 and those in (B5, B6) to be hoisted at B2. Even if instructions in (B3, B4)
could not be hoisted, for some reason, to B1, the instructions in (B5, B6) can
still be hoisted -- if legal and profitable -- to B2.

@@

% TODO: Rewrite the following.
In our current implementation we keep as many candidates in one set as possible
(greedy approach.) To start with, first two candidates are tried for
hoistability (see Section~\ref{subsec:optimistic},) if that passes then the next
instruction is tried (if available) and so on until a non-hoistable instruction
is found. The partition starts from that point.  We split the list at a point
where the legality checks fail to hoist subset of candidates which are legal to
hoist and then start finding new hoisting point for the remaining ones.

@@

The amount of hoisting depends on whether we collect \GVN{} of instructions
before finding candidates (optimistic) or, on-demand (pessimistic.) It also
depends on the generality of the \GVN{} algorithm as mentioned earlier in
Section~\ref{subsec:finding-candidates}. We have implemented an optimistic
\gcm{} of congruent instructions which uses the liveness analysis as
illustrated by \citet{das2012} and ranking expressions explained by
\citet{rosen1988global}.

\GCM{} basically consists of two parts, i.e., hoisting and sinking. This
implementation only moves congruent instructions. An immediate guarantee of this
approach is gain in code size (the final executable may be of larger size
because of more inlining.) The algorithm prefers hoisting to sinking. If the
dependency of a hoistable candidate is in the same basic-block as the candidate,
then the dependency must also be hoistable, otherwise hoisting will be illegal or
would require a complicated code generation to make it legal. The current
algorithm discards cases if the dependency is neither hoistable nor
rematerializable.

@@

Code hoisting opens new opportunities for other hoistable candidates which were
of higher rank (depended on candidates which got hoisted.) Ideally we could
iterate on lower ranking expressions first and then proceed to higher ranking
expressions in the same iteration but \LLVM{'}s \GVN{} infrastructure does not
compute equivalence classes in an effective way. We found it simpler to just
recompute the value numbers and start finding hoistable candidates again.

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that are
candidates for \gcm{}, matching the complexity of \PRE{} on \SSA{} form.
The analyses computed for this pass are Global Value Numbering, Computation of
DJ-graph and merge-sets, Marking Barriers, all are linear in number of
instructions in a function. Liveness analysis is expensive and only performed
on-demand for hoistable candidates so it does not affect the overall compilation
time by much. Other analyses like Alias Analysis, \MemorySSA{} and Dominator Tree
are already available in the \LLVM{} pass pipeline. Although we recompute \GVN{}, and
Barriers for each iteration of the \gcm{}, we have not seen significant
increase in the compilation times (see Section~\ref{tab:compile-time}.) We have also
provided appropriate compiler
flags to expedite \gcm{} by bailing out with fewer iterations, or skip the
liveness based profitability analysis to aggressively move the code as long as
they are legal.

@@

Overall there was only a minor change in the performance numbers with a few improvements 
(429.mcf, 458.sjeng, etc.) and a few regressions (447.dealII, 453.povray,
etc.) The \gcm{} pass was run twice in the pass pipeline (at -Ofast,)
first after EarlyCSE pass, and then after the inliner. Because EarlyCSE removes
local redundancies, \GCM{} would have to analyze less redundant instructions, also
as inliner creates more redundancies, \GCM{} would have more opportunities to move
congruent instructions out of branches.

We see some nice gains in code size at -Os with \gcm{} (403.gcc, 447.dealII, and
483.xalancbmk) as shown in Table~\ref{tab:code-size}.  The code size listed here
is the text size delta (from linux size command) of the final executable for
each benchmark.  On the other hand, some benchmarks like 473.astar increased in
code size by almost 2\%. This is because once the code-size of a function
decreases (due to \GCM{},) it becomes cheaper to inline. As we can see in
Table~\ref{tab:stats}, four more functions got inlined in 473.astar. Except for
447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got
more functions inlined. Also, since the pass runs early, it affects many
optimizations which rely on the number of instructions, length of the use-def
chain, and other metrics.

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|}
      \hline
 \GCM{}  &\multicolumn{2}{|c|}{Total Instructions} &\multicolumn{2}{|c|}{Scalars}  & \multicolumn{2}{|c|}{Loads} & \multicolumn{1}{|c|}{Stores} &Calls   &Instructions    \\\hline
 metric           &  H          &R        &H      &R       &H           &R              &H=R    &H & sunk \\\hline
 400.perlbench	  &  2387	& 2456	  &  855  & 857	   & 1493	& 1560		& 38	& 1	& 4	        \\\hline
 401.bzip2	  &  176	& 300	  &  27	  & 27	   & 149	& 273		& 0	& 0 	& 0		\\\hline
 403.gcc	  &  5670	& 5761	  &  2558 & 2567   & 3109	& 3191		& 3	& 0	& 1	        \\\hline
 429.mcf	  &  42	        & 43	  &  5	  & 5	   & 37 	& 38		& 0	& 0 	& 0		\\\hline
 433.milc	  &  128	& 130	  &  10	  & 10	   & 117	& 119		& 0	& 1	& 0		\\\hline
 444.namd	  &  262	& 264	  &  113  & 113	   & 148	& 150		& 0	& 1	& 0		\\\hline
 445.gobmk	  &  1021	& 1052	  &  312  & 314	   & 705	& 734		& 3	& 1	& 1	        \\\hline
 447.dealII	  &  10455	& 12325	  &  4009 & 4063   & 6394	& 8210		& 20	& 32	& 0	        \\\hline
 450.soplex	  &  2389	& 2731	  &  927  & 1004   & 1437	& 1702		& 14	& 11	& 0	        \\\hline
 453.povray	  &  1967	& 2269	  &  560  & 590	   & 1389	& 1661		& 4	& 14	& 0	        \\\hline
 456.hmmer	  &  408	& 427	  &  82	  & 82	   & 319	& 338		& 1	& 6	& 0	        \\\hline
 458.sjeng	  &  372	& 399	  &  128  & 138	   & 234	& 251		& 1	& 9	& 1	        \\\hline
 462.libquantum	  &  49	        & 49	  &  17	  & 17	   & 32	      	& 32		& 0	& 0	& 0		\\\hline
 464.h264ref	  &  2953	& 3438	  &  712  & 779	   & 2238	& 2656		& 2	& 1	& 1	        \\\hline
 470.lbm	  &  23	        & 23	  &  4	  & 4	   & 19		& 19		& 0	& 0	& 0		\\\hline
 471.omnetpp	  &  391	& 408	  &  92	  & 92	   & 297	& 314		& 0	& 2	& 0		\\\hline
 473.astar	  &  202	& 222	  &  62	  & 64	   & 140	& 158		& 0	& 0	& 0		\\\hline
 482.sphinx3	  &  168	& 172	  &  32	  & 32	   & 136	& 140		& 0	& 0	& 0		\\\hline
 483.xalancbmk	  &  5894	& 6162	  &  2586 & 2660   & 3301	& 3495		& 4	& 3	& 2	        \\\hline
 Grand total      &34957	&38631	  &13091  &13418   &21694	&25041	        &90	&82     &10             \\\hline
    \end{tabular}
  \end{center}
  \caption{\GCM{} stats on SPEC2006 at -Ofast, H(Hoisted,) R(Removed); Loads are
    hoisted the most followed by scalars, stores and calls.}
  \label{tab:code-motion-metric}
\end{table}
